Nanocall: An Open Source Basecaller for Oxford Nanopore Sequencing Data

### 2016-03-17 paper-biorxiv/paper.tex
more pore models edits
strand seperation

Nanocall is designed to work with 6-mer pore models. By default we use the models provided by Metrichor. These are 3 models, 1 for the template strand and 2 for the complement strand. Emission probabilities are calculated by multiplying the probability density from the Gaussian and Inverse Gaussian distributions that model the event mean and event standard deviation, respectively. This is done after the models are scaled (see Scaling below.)

@@

In the first step, we use a heuristic to detect the \emph{abasic current level}. In general, the currents corresponding to regular DNA are in the range 50-90 pA, and those corresponding to abasic DNA are higher than 100 pA. However, the exact levels are affected by the \texttt{shift} and \texttt{scale} parameters, which are not known a priori. For that reason, Nanocall uses a heuristic to estimate abasic current level, specifically: take the current level larger than 99\% of all currents in that read, and add 5pA.

In the second step, Nanocall looks for islands of 5 or more consecutive abasic current measurements. Then, islands within 50bp of each other are merged. Next, Nanocall selects the island closest to the middle of the event sequence. If this island is within the middle third of the entire event sequence, then it is used to separate the events corresponding to the 2 strands. If on the other hand this island is outside of the middle third of the event sequence, Nanocall gives up trying to separate the strands and attempts to basecall the entire event sequence as a template strand.

>>>>

Nanocall is designed to work with 6-mer pore models. By default we use the models provided by Metrichor. Metrichor provides three models, one for the template strand and two for the complement strand. Emission probabilities are calculated by multiplying the probability density from the Gaussian and Inverse Gaussian distributions that model the event mean and event standard deviation, respectively. This is done after the models are scaled (see Scaling below.)

@@

In the first step, we use a heuristic to detect the \emph{abasic current level}. In general, the currents corresponding to regular DNA are in the range 50-90 pA, and those corresponding to abasic DNA are higher than 100 pA. However, the exact levels are affected by the \texttt{shift} and \texttt{scale} parameters, which are not known a priori. For that reason, Nanocall uses a heuristic to estimate abasic current level. Specifically we take the current level larger than 99\% of all event levels in the read, and add 5pA.

In the second step, Nanocall looks for islands of 5 or more consecutive abasic current measurements. Then, islands within 50bp of each other are merged. Next, Nanocall selects the island closest to the middle of the event sequence. If this island is within the middle third of the entire event sequence it is used to separate the events corresponding to the two strands. If on the other hand this island is outside of the middle third of the event sequence, Nanocall gives up trying to separate the strands and attempts to basecall the entire event sequence as a template strand.
